---
title: "Comorbidity_Mapping_Demo"
author: "Meg Hutch, Trang Le"
date: "1/15/2021"
output: html_document
---


```{r message=FALSE, warning=FALSE, include=FALSE}
my_packages <- c('tidyverse', 'readxl', 'DT', 'rcartocolor', 'gghighlight', 
                 'cowplot', 'broom', 'devtools', 'tableone')

existing_pkgs <- installed.packages()[,"Package"]
to_install <- setdiff(my_packages, existing_pkgs)
if (length(to_install)) install.packages(to_install)
if (!'icd' %in% existing_pkgs) devtools::install_github('jackwasey/icd')

lapply(my_packages, library, character.only = TRUE)
theme_set(theme_bw() + 
            theme(legend.title = element_blank(),
                  panel.grid.minor = element_blank()))

source("R/comorbidity-map-funcs.R")
```

## Import Data

Replace path of simulated_data with that of your site's Phase2.1 LocalPatientObservations csv file.
```{r}
data <- read.csv("data/simulated_data.csv")
```

# Elixhauser Comorbidity Mapping

## Map Elixhauser Comorbidities

t1: earliest time point to consider comorbidities
t2: latest time point to consider comorbidities
example <- t1 = -365, and t2 = -1 will map all all codes up to a year prior but before admission (admission = day 0)

```{r}
comorb_names_elix <- get_quan_elix_names()
comorbs_elix <- as.vector(comorb_names_elix$Abbreviation)

comorb_elix <- map_char_elix_codes(
  df = data,
  comorb_names = comorb_names_elix,
  t1 = -365,
  t2 = -15,
  map_type = 'elixhauser',
  truncate = TRUE
)
```

1- Dataframe of the Charlson/Elixhauser comorbidity index scores

2- Dataframe of the specific mapped codes arranged by frequency.
Comorbidity mapping of all patients with ICD codes prior to admission, regardless of whether or not they had a comorbidity.

```{r}
index_scores_elix <- comorb_elix$index_scores %>% 
  rename('elixhauser_score' = van_walraven_score)

# van Walraven is a modification of Elixhauser comorbidity measure
# doi.org/10.1097/MLR.0b013e31819432e5
mapped_codes_table_elix <- comorb_elix$mapped_codes_table

comorb_names_elix$Abbreviation <- as.character(comorb_names_elix$Abbreviation)
```


### Patient Level Comorbidity Scores

```{r}
index_scores_elix %>% 
  datatable() 
```

```{r}
ggplot(index_scores_elix, aes(x = elixhauser_score)) + 
  geom_histogram(bins = 15, fill = 'slateblue')
```


### Number of patients with specific comorbidities

```{r}
mapped_codes_table_elix %>%
  datatable() 
```

# Charlson Comorbidity Mapping

## Map Charlson Comorbidities

t1: earliest time point to consider comorbidities
t2: latest time point to consider comorbidities
example <- t1 = -365, and t2 = -1 will map all all codes up to a year prior but before admission (admission = day 0)

```{r}
comorb_names_char <- get_charlson_names()
comorbs_char <- as.vector(comorb_names_char$Abbreviation)

comorb_char <- map_char_elix_codes(
  df = data,
  comorb_names = comorb_names_char,
  t1 = -365,
  t2 = -15,
  map_type = 'charlson',
  truncate = TRUE
)
```

1- Dataframe of the Charlson/Elixhauser comorbidity index scores

2- Dataframe of the specific mapped codes arranged by frequency.
Comorbidity mapping of all patients with ICD codes prior to admission, regardless of whether or not they had a comorbidity.

```{r}
index_scores_char <- comorb_char$index_scores

mapped_codes_table_char <- comorb_char$mapped_codes_table

comorb_names_char$Abbreviation <- as.character(comorb_names_char$Abbreviation)
```


### Patient Level Comorbidity Scores

```{r}
index_scores_char %>% 
  datatable() 
```

```{r}
ggplot(index_scores_char, aes(x = charlson_score)) + 
  geom_histogram(bins = 15, fill = 'slateblue')
```

### Number of patients with specific comorbidities

```{r}
mapped_codes_table_char %>%
  datatable() 
```

